---
title: "Human Activity Recognition"
author: "Sergey Cherkasov"
date: "14 Oct 2015"
output: html_document
---

#Synopsis 
 
Activity trackers make possible to collect a large amount of data about personal activity. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it.
 
In this paper we are going to build a model that can predict the manner in which people did the exercise. We use  dataset, generously provided by Groupware@LES in [Qualitative Activity Recognition of Weight Lifting Exercises](http://groupware.les.inf.puc-rio.br/work.jsf?p1=11201).

#Exploratory analysis 

a 

#Building the model 

a:
```{r echo=FALSE}
###a
```

#Cross validation 

a

```{r echo=FALSE}
###a
```


#out of sample error 

a

```{r echo=FALSE}
###a
```


#Conclusion 
 
a

#Appendices

##Appendix 1. .
```{r echo=FALSE}
###a
```

##Appendix 2. .
``` {r echo=FALSE}
###a
```

##Appendix 3. .
``` {r echo=FALSE}
### a
```

##Appendix 4. .
``` {r echo=FALSE}
### a
```

##Appendix 5. .
``` {r echo=FALSE}
### a
```